Testing the motor and cognitive foundations of Paleolithic social transmission

### 2021-08-17 manuscript/manuscript.Rmd
figures remade using R.

4)  Fluid intelligence [@cattell1963] refers to the capacity to engage in abstract reasoning and problem solving in a way that is minimally dependent on prior experience. It complements "crystallized intelligence" (the ability to apply learned procedures and knowledge) as one of the two factors (gf, gc) comprising so-called "general intelligence" (g). Fluid intelligence is closely related to the executive control of attention and manipulation of information held in working memory [@engle2018](Engle 2018). It is hypothesized to support technological innovation [@coolidge2005] and/or the intentional learning of new skills [@unsworth2005; @stout2015]. We measured fluid intelligence using the short version [@bilker2012] of the classic Raven Progressive Matrices task, which requires participants to complete increasingly difficult pattern matching questions.

@@

While we did not find significant differences in central tendency between our two experimental groups, results (**Figure** \@ref(fig:fig7)) did indicate lower variance in the trained group. To test whether training reduced variability in performance outcomes between subjects, we compared variance metrics between the trained and untrained individuals using the F-test on either core-averaged or flake specific variances. **Table** \@ref(tab:tab5) and **Figure** \@ref(fig:fig9) present the results from these comparisons showing significant variance differences predominantly in flaking Quality, number of large DPs, core delta mass, and total amount of flaked mass). In most instances, variance in the untrained group exceeds that of trained individuals by 1.5 to 4.7 times. The most salient effect of instruction was thus not to shift mean performance but to reduce variability by eliminating the skew (generally toward poorer outcomes) seen in the untrained group (**Figure** \@ref(fig:fig9)), rather than to shift the mean.
```{r tab5, echo=FALSE, message=FALSE}
library(tidyverse)
df5<-read.csv(here::here("data/tables/table5.csv"), fileEncoding = 'UTF-8-BOM')
kableExtra::kbl(df5, booktabs = T, caption = "F-test results comparing variances between trained and untrained subjects across 
our various flaking performance metrics.") %>%
  kableExtra::kable_styling(latex_options = c("striped", "HOLD_position"), font_size = 7)%>%
kableExtra::row_spec(0, bold = T)
```

@@

**Table** \@ref(tab:tab7) presents this model's coefficients and summary outputs, wherein baseline refers to the untrained condition with all continuous predictors at the sample average. The parameter estimates for the continuous predictors reflect the expected change in utility for 1 standard deviation change in the predictor variable. We found significant (p\< 0.05) and substantial (Standardized Estimate \>= 0.50, i.e. a 50% change in variable) main effects of Grip Strength, Visuospatial nBack, and BEAST. The main effect of Grip Strength (**Figure** \@ref(fig:fig12)), irrespective of learning condition, indicates the basic importance of strength in generating higher production rates among naive knappers at least when efficiency and quality are not considered.
```{r fig12, echo=FALSE, out.width = '100%', fig.align='center', fig.cap="Significant main effect of grip strength on overall flaking productivity."}
knitr::include_graphics(here::here("figure/Figure12.jpeg"))
```

>>>>

4)  Fluid intelligence [@cattell1963] refers to the capacity to engage in abstract reasoning and problem solving in a way that is minimally dependent on prior experience. It complements "crystallized intelligence" (the ability to apply learned procedures and knowledge) as one of the two factors ($g_f$, $g_c$) comprising so-called "general intelligence" (g). Fluid intelligence is closely related to the executive control of attention and manipulation of information held in working memory [@engle2018]. It is hypothesized to support technological innovation [@coolidge2005] and/or the intentional learning of new skills [@unsworth2005; @stout2015]. We measured fluid intelligence using the short version [@bilker2012] of the classic Raven Progressive Matrices task, which requires participants to complete increasingly difficult pattern matching questions.

@@

While we did not find significant differences in central tendency between our two experimental groups, results (**Figure** \@ref(fig:fig7)) did indicate lower variance in the trained group. To test whether training reduced variability in performance outcomes between subjects, we compared variance metrics between the trained and untrained individuals using the F-test on either core-averaged or flake specific variances. **Table** \@ref(tab:tab5) and **Figure** \@ref(fig:fig9) present the results from these comparisons showing significant variance differences predominantly in flaking Quality, number of large DPs, core delta mass, and total amount of flaked mass). In most instances, variance in the untrained group exceeds that of trained individuals by 1.5 to 4.7 times. The most salient effect of instruction was thus not to shift mean performance but to reduce variability by eliminating the skew (generally toward poorer outcomes) seen in the untrained group (**Figure** \@ref(fig:fig9)), rather than to shift the mean.
```{r tab5, echo=FALSE, message=FALSE}
library(tidyverse)
df5<-read.csv(here::here("data/tables/table5.csv"), fileEncoding = 'UTF-8-BOM')
colnames(df5)[colnames(df5) %in% c("X95..CI")] <- c("95% CI")
kableExtra::kbl(df5, booktabs = T, caption = "F-test results comparing variances between trained and untrained subjects across 
our various flaking performance metrics.") %>%
  kableExtra::kable_styling(latex_options = c("striped", "HOLD_position"), font_size = 7)%>%
kableExtra::row_spec(0, bold = T)
```

@@

**Table** \@ref(tab:tab7) presents this model's coefficients and summary outputs, wherein baseline refers to the untrained condition with all continuous predictors at the sample average. The parameter estimates for the continuous predictors reflect the expected change in utility for 1 standard deviation change in the predictor variable. We found significant (p\< 0.05) and substantial (Standardized Estimate \>= 0.50, i.e. a 50% change in variable) main effects of Grip Strength, Visuospatial nBack, and BEAST. The main effect of Grip Strength (**Figure** \@ref(fig:fig12)), irrespective of learning condition, indicates the basic importance of strength in generating higher production rates among naive knappers at least when efficiency and quality are not considered.
```{r fig12, echo=FALSE, out.width = '100%', fig.align='center', fig.cap="Significant main effect of grip strength on overall flaking productivity."}
knitr::include_graphics(here::here("figure/Figure12.jpg"))
```
